---
title: "point pattern analysis"
author: "Ying Chen"
date: "29/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(sf)
library(tidyverse)
library(dplyr)
library(janitor)
library(RColorBrewer)
library(classInt)
library(sp)
library(rgeos)
library(rgdal)
library(spatstat)
library(here)
library(maptools)
library(GISTools)
library(tmap)
library(geojson)
library(geojsonio)
library(tmaptools)
library(hexbin)
library(ggspatial)
library(ggsn)
library(raster)
library(fpc)
library(dbscan)
library(plotrix)
library(spdep)
library(ggplot2)
library(ggpubr)
library(factoextra) 
library(knitr)
library(OpenStreetMap)
library(viridis)
```

## Data Preparation

```{r message=FALSE}
boroMap <- st_read(here::here("data","statistical-gis-boundaries-london", "ESRI",
                              "London_Borough_Excluding_MHW.shp"))%>%
  st_transform(., 27700)

wardMap <- st_read(here::here("data","statistical-gis-boundaries-london", "ESRI",
                              "London_Ward_CityMerged.shp")) %>% 
  st_transform(., 27700)

msoaMap <- st_read(here::here("data","statistical-gis-boundaries-london", "ESRI",
                              "MSOA_2011_London_gen_MHW.shp")) %>% 
  st_transform(., 27700)
```

The following function used to search files that follow a specific pattern is obtained from <https://github.com/sjaraha/clustering-spatiotemporal-data>.

```{r}
list_data_paths <- function(pattern, rec){
  # searches working directory for files that match the specified pattern
  # on match, adds file path to a list
  # returns list the list of matching file paths
  ## pattern (str): regex pattern to match
  ## rec (boolean): recurse into directories (True) or don't (False)
  
  # initialize list
  data_path_list <- c()
  # loop through directories
  for (pd in list.dirs(recursive = rec)){
    # loop through files in directories
    for (f in list.files(pd)){
      # find files that match the pattern
      if (grepl(pattern, f, ignore.case = FALSE)==TRUE){
        # construct path to matching file
        data_path <- paste(pd,f, sep="/")
        # add path to list
        data_path_list <- c(data_path_list,data_path)
      }}}
  # return list of paths to matching files
  return(data_path_list)
}
```

Use the function above and search for all crime data (ended with metropolitan-street.csv) in my local folder.

```{r message = FALSE, warning=FALSE}
setwd("/Users/yingchen/Documents2/CASA/dissertation/dissertation2021/data/crime_data")
crime_list <- list_data_paths("\\-metropolitan-street.csv$",FALSE) 

crime <- crime_list %>% 
  lapply(read_csv) %>% 
  bind_rows
```

Remove duplicated rows, clean names and select points within London.

```{r message = FALSE}
crime<- crime %>% 
  clean_names() %>% 
  distinct(.) %>% 
  filter(latitude != "NA" | longitude != "NA") %>% 
  st_as_sf(., coords = c("longitude", "latitude"), 
           crs = 4326) %>% 
  st_transform(., 27700) 

crime_london <- crime[boroMap,]
```

Search for stop and search data (ends with -metropolitan-stop-and-search.csv).

```{r echo=TRUE, message = FALSE}
setwd("/Users/yingchen/Documents2/CASA/dissertation/dissertation2021/data/crime_data")
ss_list <- list_data_paths("\\-metropolitan-stop-and-search.csv$",FALSE) 

ss <- ss_list %>% 
  lapply(read_csv) %>% 
  bind_rows
```

clean names, remove duplicated records and select points within London boroughs.

```{r message=FALSE, warning=False}
ss <- ss %>% 
  clean_names() %>% 
  distinct(.) %>% 
  filter(latitude != "NA" | longitude != "NA") %>% 
  st_as_sf(., coords = c("longitude", "latitude"), 
           crs = 4326) %>% 
  st_transform(., 27700) 

ss_london <- ss[boroMap,]
```

prepare for the basemap setting

```{r message=FALSE, warning=FALSE}
# create a bounding box of London
LondonBB <- boroMap %>%
  st_transform(., 4326)%>%
  st_bbox()

library(OpenStreetMap)
# set the basemap showing London 
basemap <- OpenStreetMap::openmap(c(51.2867602,-0.5103751),c(51.6918741,0.3340156), 
                                  zoom=NULL,
                                  "stamen-toner")

# convert the basemap to British National Grid
basemap_bng <- openproj(basemap, projection="+init=epsg:27700")
```

Function to extract coordinates

```{r}
# before plotting, we need to transform geometry to (x,y) coordinates and the function was found at <https://maczokni.github.io/crimemapping_textbook_bookdown/more-on-thematic-maps.html>
sfc_as_cols <- function(x, names = c("x","y")) {
  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
  ret <- sf::st_coordinates(x)
  ret <- tibble::as_tibble(ret)
  stopifnot(length(names) == ncol(ret))
  x <- x[ , !names(x) %in% names]
  ret <- setNames(ret,names)
  dplyr::bind_cols(x,ret)
}
```

## dbscan

DBSCAN requires two parameters: 1. Epsilon - the radius within which the algorithm with search for clusters 2. MinPts - the minimum number of points that should be considered a cluster.

Two ways to determine Epsilon: 1. Ripley's K - where K values above the Poisson distribution indicates areas of clustering; needs to find the K values' cutoff between above and below the Poisson distribution. 2. KNNDistance - it is the distance of each point to its k-th nearest neighbor; needs to find a knee in the plot. "The idea behind this heuristic is that points located inside of clusters will have a small k-nearest neighbor distance, because they are close to other points in the same cluster, while noise points are more isolated and will have a rather large kNN distance (Hahsler et al., 2019).

Potential way to determine MinPts: The rule of thumb for setting minPts is to use at least the number of dimensions of the dataset plus one (Hahsler et al., 2019). However, by setting MinPts = 3 (which is what Hahsler et al suggest) it results in 229 unique cluster for all S&S points, which is too many.

### Black arrests

```{r message=FALSE, warning=FALSE}
window <- as.owin(msoaMap)

ss_black <- ss_london %>% 
  filter(officer_defined_ethnicity == "Black") %>% 
  filter(outcome == "Arrest")

ss_blackSP <- ss_black %>%  
  as(., 'Spatial') 
 
ss_blackPPP <- ppp(x=ss_blackSP@coords[,1],
                     y=ss_blackSP@coords[,2],
                     window=window)

ss_blackPoints<- ss_blackSP %>%
  geometry(.)%>%
  as.data.frame()
```

1.  Choose Ripley's K method to select suitable parameters:

```{r message=FALSE, warning=FALSE}
png("output/Rip_Black.png", units="in", width=6, height=4, res=300)
r1 <- ss_blackPPP %>%
  Kest(., correction="border") %>%
  plot(xlim=c(0,1000),
       main="Ripley's K Test \n Black S&S Arrests (threshold = 400)") %>% 
  abline(v = 400, col = "blue", lty = 2)
dev.off()
```

Distance = 400, minimum points = 50

```{r warning=F, message=F}
library(factoextra)

db_b <- ss_blackPoints %>%
  fpc::dbscan(.,eps = 400, MinPts = 50)

theme_set(theme_minimal())

fviz_cluster(db_b, ss_blackPoints, 
                           geom = "point",
                           stand = FALSE, labelsize = NA,
                           outlier.pointsize = .8,
                           #xlab="x", ylab="y",
                           main="DBSCAN Cluster Hulls \n") +
theme(plot.title=element_text(size = 12, hjust = 0.5, face="bold"),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
  coord_equal()

```

2.  Choose KNN distance method to select suitable parameters:

```{r}
ss_blackPoints %>%
  dbscan::kNNdistplot(.,k=4) %>% 
  title(main="4-nearst Neighbor Distance Plot") %>% 
  abline(h = 500, col = "red", lty = 2)
```

KNN and Ripley's K yields a similar result

```{r message=FALSE, warning=FALSE}
db_b <- ss_blackPoints %>%
  dbscan(., eps= 400, minPts = 50)

library(OpenStreetMap)

# add (x,y) coordinate and clusters to new columns  
ss_black <- sfc_as_cols(ss_black, c("longitude", "latitude")) %>% 
  mutate(dbcluster=db_b$cluster)  

# create convex hull polygons to wrap around the points in our clusters
hulls_b <- ss_black %>%
  group_by(dbcluster) %>%
  dplyr::mutate(hull = 1:n(),
                hull = factor(hull, chull(longitude, latitude)))%>%
  arrange(hull) %>% 
  filter(dbcluster >=1) # since dbcluster < 1 are noises


# plot

autoplot.OpenStreetMap(basemap_bng) + 
  geom_point(data=hulls_b,
             aes(longitude,latitude,
                 colour=dbcluster),
             size=0.1, alpha=0.3) + 
    scale_color_viridis(option = "D",
                        begin = 0.9,
                        end = 0.5, 
                        name = "Black Arrests",
                        labels = NULL) +
  theme(legend.position = "right") +
  ggtitle("DBSCAN Cluster for S&S Only Targeted to Black People")+
  theme(plot.title=element_text(hjust = 0.5, face="bold"))+
  labs(caption = "Copyright OpenStreetMap contributors",
       x="Meter",y="Meter")
```

### White arrests

```{r message=FALSE, warning=FALSE}
ss_white<- ss_london %>% 
  filter(outcome == "Arrest") %>% 
  filter(officer_defined_ethnicity == "White") 

ss_whiteSP <- ss_white %>%  
  as(., 'Spatial') 

ss_whitePPP <- ppp(x=ss_whiteSP@coords[,1],
                     y=ss_whiteSP@coords[,2],
                     window=window)

ss_whitePoints<- ss_whiteSP %>%
  geometry(.)%>%
  as.data.frame()
```

1.  Ripley's K method:

```{r message=FALSE, warning=FALSE}
png("output/Rip_White.png", units="in", width=6, height=4, res=300)
r1 <- ss_whitePPP %>%
  Kest(., correction="border") %>%
  plot(xlim=c(0,1000),
       main="Ripley's K Test \n White S&S Arrests (threshold = 450)") %>% 
  abline(v = 450, col = "blue", lty = 2)
dev.off()
```

```{r message=FALSE, warning=FALSE}
db_w <- ss_whitePoints %>%
  fpc::dbscan(.,eps = 450, MinPts = 50)

theme_set(theme_minimal())

fviz_cluster(db_w, ss_whitePoints, 
                           geom = "point",
                           stand = FALSE, labelsize = NA,
                           outlier.pointsize = .8,
                           #xlab="x", ylab="y",
                           main="DBSCAN Cluster Hulls") +
theme(plot.title=element_text(size = 12, hjust = 0.5, face="bold"),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
  coord_equal()
```

```{r message=FALSE, warning=FALSE}
ss_whitePoints %>%
  dbscan::kNNdistplot(.,k=4) %>% 
  title(main="4-nearst Neighbor Distance Plot") %>% 
  abline(h = 500, col = "red", lty = 2)
```

```{r message=FALSE, warning=FALSE}
db_w <- ss_whitePoints %>%
  fpc::dbscan(.,eps = 450, MinPts = 50)

# add (x,y) coordinate and clusters to new columns  
ss_white <- sfc_as_cols(ss_white, c("longitude", "latitude")) %>% 
  mutate(dbcluster=db_w$cluster)  

# create convex hull polygons to wrap around the points in our clusters
hulls_w <- ss_white %>%
  group_by(dbcluster) %>%
  dplyr::mutate(hull = 1:n(),
                hull = factor(hull, chull(longitude, latitude)))%>%
  arrange(hull) %>% 
  filter(dbcluster >=1) # since dbcluster < 1 are noises

# plot
autoplot.OpenStreetMap(basemap_bng) + 
  geom_point(data=hulls_w,
             aes(longitude,latitude,
                 colour=dbcluster),
             size=0.1, alpha=0.3) + 
    scale_color_gradient(low = "cadetblue4", 
                          high = "cadetblue4", 
                          na.value = NA,
                          name = "White Arrests",
                          labels = NULL)+
  theme(legend.position = "right") +
  ggtitle("DBSCAN Cluster for S&S Only Targeted to White People")+
  theme(plot.title=element_text(hjust = 0.5, face="bold"))+
  labs(caption = "Copyright OpenStreetMap contributors",
       x="Meter",y="Meter")

autoplot.OpenStreetMap(basemap_bng) + 
  geom_point(data=hulls_b,
             aes(longitude,latitude,
                 colour=dbcluster),
             size=0.1, alpha=0.3) + 
    scale_color_viridis(option = "D",
                        begin = 0.9,
                        end = 0.5, 
                        name = "Black Arrests",
                        labels = NULL) +
  theme(legend.position = "right") +
  ggtitle("DBSCAN Cluster for S&S Only Targeted to Black People")+
  theme(plot.title=element_text(hjust = 0.5, face="bold"))+
  labs(caption = "Copyright OpenStreetMap contributors",
       x="Meter",y="Meter")
```

### Asian

```{r message=FALSE, warning=FALSE}
ss_asian <- ss_london %>% 
  filter(outcome == "Arrest") %>% 
  filter(officer_defined_ethnicity == "Asian") 

ss_asianSP <- ss_asian %>% 
  as(., 'Spatial') 

ss_asianPPP <- ppp(x=ss_asianSP@coords[,1],
                     y=ss_asianSP@coords[,2],
                     window=window)

ss_asianPoints<- ss_asianSP %>%
  geometry(.)%>%
  as.data.frame()
```

1.  Ripley's K:

```{r message=FALSE, warning=FALSE}
png("output/Rip_Asian.png", units="in", width=6, height=4, res=300)
r1 <- ss_asianPPP %>%
  Kest(., correction="border") %>%
  plot(xlim=c(0,1000),
       main="Ripley's K Test \n Asian S&S Arrests (threshold = 550)") %>% 
  abline(v = 550, col = "blue", lty = 2)
dev.off()
```

```{r message=FALSE, warning=FALSE}
db_a <- ss_asianPoints %>%
  fpc::dbscan(.,eps = 550, MinPts = 50)

theme_set(theme_minimal())

fviz_cluster(db_a, ss_asianPoints, 
                           geom = "point",
                           stand = FALSE, labelsize = NA,
                           outlier.pointsize = .8,
                           #xlab="x", ylab="y",
                           main="DBSCAN Cluster Hulls") +
theme(plot.title=element_text(size = 12, hjust = 0.5, face="bold"),
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
  coord_equal()

```

```{r message=FALSE, warning=FALSE}
ss_asianPoints %>%
  dbscan::kNNdistplot(.,k=4) %>% 
  title(main="4-nearst Neighbor Distance Plot") %>% 
  abline(h = 1000, col = "red", lty = 2)
```

```{r message=FALSE, warning=FALSE}
db_a <- ss_asianPoints %>%
  fpc::dbscan(.,eps = 550, MinPts = 50)

# add (x,y) coordinate and clusters to new columns  
ss_asian <- sfc_as_cols(ss_asian, c("longitude", "latitude")) %>% 
  mutate(dbcluster=db_a$cluster)  

# create convex hull polygons to wrap around the points in our clusters
hulls_a <- ss_asian %>%
  group_by(dbcluster) %>%
  dplyr::mutate(hull = 1:n(),
                hull = factor(hull, chull(longitude, latitude)))%>%
  arrange(hull) %>% 
  filter(dbcluster >=1) # since dbcluster < 1 are noises

# plot
autoplot.OpenStreetMap(basemap_bng) + 
  geom_point(data=hulls_a,
             aes(longitude,latitude,
                 colour=dbcluster),
             size=0.1, alpha=0.3) + 
    scale_color_gradient(low = "gold2", 
                          high = "gold2", 
                          na.value = NA,
                          name = "Asian Arrests",
                          labels = NULL)+
  theme(legend.position = "right") +
  ggtitle("(c)")+
  theme(plot.title=element_text(hjust = 0.05, face="bold"))+
  labs(caption = "Copyright OpenStreetMap contributors",
       x="Meter",y="Meter")
```

### Overlapping map

```{r}
a <- autoplot.OpenStreetMap(basemap_bng) +
  geom_point(data=hulls_b,
             aes(longitude,latitude,
                 colour=dbcluster),
             size=0.1, alpha=0.3) + 
    scale_color_viridis(option = "D",
                        begin = 0,
                        end = 0, 
                        name = "Black \nArrests",
                        labels = NULL) +
  labs(title = "DBSCAN, Epsilon=400 m, MinPts=50",
       x="Meter", y="Meter") +
 theme(axis.text=element_text(size=7),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size= 7),
       legend.position = "right",
       plot.title = element_text(face = "bold", size = 10)) 

b <- autoplot.OpenStreetMap(basemap_bng) + 
  geom_point(data=hulls_w,
             aes(longitude,latitude,
                 colour=dbcluster),
             size=0.1, alpha=0.3) + 
    scale_color_viridis(option = "D",
                        begin = 0.5,
                        end = 0.5, 
                        name = "White \nArrests",
                        labels = NULL) +
  ggtitle("DBSCAN, Epsilon=450 m, MinPts=50")+
 labs(x="Meter", y="Meter") +
 theme(axis.text=element_text(size=7),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size= 7),
       legend.position = "right",
       plot.title = element_text(face = "bold", size = 10)) 

c <-  autoplot.OpenStreetMap(basemap_bng) + 
  geom_point(data=hulls_a,
             aes(longitude,latitude,
                 colour=dbcluster),
             size=0.1, alpha=0.3) + 
    scale_color_viridis(option = "D",
                        begin = 1,
                        end = 1, 
                        name = "Asian \nArrests",
                        labels = NULL) +
    ggtitle("DBSCAN, Epsilon=550 m, MinPts=50")+
   labs(x="Meter", y="Meter") +
 theme(axis.text=element_text(size=7),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size= 7),
       legend.position = "right",
       plot.title = element_text(face = "bold", size = 10)) 

d <- autoplot.OpenStreetMap(basemap_bng) +
  geom_point(data=hulls_b, 
             aes(longitude,latitude),
             size=0.1, alpha=0.3) +
  geom_polygon(data = hulls_b,
               aes(longitude,latitude,
                   group=dbcluster,
                   fill=dbcluster),
               fill="#440154FF") +
  geom_point(data=hulls_w, 
             aes(longitude,latitude),
             size=0.1, alpha=0.3)+
  geom_polygon(data = hulls_w, 
               aes(longitude,latitude, 
                   group=dbcluster,
                   fill=dbcluster),
               fill = "#1F968BFF") +
  geom_point(data=hulls_a, 
             aes(longitude,latitude), 
             size=0.1, alpha=0.3)+
  geom_polygon(data = hulls_a, 
               aes(longitude,latitude, 
                   group=dbcluster,
                   fill=dbcluster),
               fill = "#FDE725FF") +
  labs(caption = "Copyright OpenStreetMap contributors",
       x="Meter",y="Meter") +
 theme(axis.text=element_text(size=7),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size= 7),
       legend.position = "right") 

figure <- ggarrange(a, b, c, d + rremove("x.text"), 
          labels = c("(a)", "(b)", "(c)", "(d)"),
          font.label = (size = 10),
          ncol = 2, nrow = 2) +
  theme(plot.margin = margin(0.1,0.1,2,0.1, "cm")) 

  
ggsave(file="output/dbscan_final.png", figure)

figure
```

## st-dbscan

The 'stdbscanr' package and related functions was obtained from Dr.Gordon McDonald's github at <https://github.com/gdmcdonald/stdbscanr>

```{r message=FALSE}
# install.packages("devtools")
# devtools::install_github("gdmcdonald/stdbscanr")
# install.packages("data.table")          

library("data.table")
library(stdbscanr)

ss_londonSub <- ss_london[sample(nrow(ss_london), 5000), ] %>%  # random select 100 rows
  st_transform(., 4326) %>% 
  sfc_as_cols(., c("longitude", "latitude")) %>%    # convert geometry to coordinates
  setDT(.) %>%     # convert to data.table
  setkey(., date)  # sort by data 

# add time intervals in minues between points
ss_londonSub[,time_inc := as.numeric(date - shift(date), units = "mins")]

# run st-dbscan
location_with_visits <- 
  get_clusters_from_data(df = ss_londonSub,
                         x = "longitude", 
                         y = "latitude", 
                         t = "date",
                         eps = 0.010,  # 0.005 latitude/longitude ~ 500m either way in London
                         eps_t = 1440, # 1440 minutes = 1day
                         minpts = 10)
```

```{r results='asis'}
#Define a mode function to get the most common label
mode <- function(x) { names(which.max(table(x))) }

#data.table summary table of visits
clusters <-                                   
  location_with_visits[     
    !is.na(cluster),                         
    .(n = .N,                               
      latitude = mean(latitude),            
      longitude = mean(longitude), 
      time_spent = sum(time_inc,na.rm = T),
      ethnicity_label = mode(officer_defined_ethnicity)            
    ),    
    by=cluster]

stats <- clusters %>% 
  arrange(desc(n)) %>% 
  rename(number_searchs = "n")

kable(stats, caption = "ST-DBSCAN Cluster Labels")
```

```{r}
stats %>% 
  group_by(ethnicity_label) %>% 
  summarise(., count=n()) %>% 
  arrange(desc(count)) %>% 
  kable()
```

```{r}
stats %>% 
  group_by(ethnicity_label, time_spent) %>% 
  group_by(ethnicity_label) %>% 
  summarise(., time=sum(time_spent)) %>% 
  arrange(desc(time)) %>% 
  kable()
```


```{r message=FALSE}
# order by time
setkey(location_with_visits, date)

library(leaflet)
library(htmlwidgets)
# plot on leaflet map
m <- leaflet(data = clusters) %>% 
  addProviderTiles("OpenStreetMap.Mapnik") %>% 
  addMarkers(popup = ~paste0("Time spent: ",round(time_spent/60, 1), " hours.<br>",
                             "S&S Ethnicity Cluster ",cluster,": ",ethnicity_label,"<br>",
                             "Counts: ", n))
m

saveWidget(m, file="stdbscan_map.html", selfcontained = TRUE)
```

```{r}
stPlot <- location_with_visits %>% 
  st_as_sf(.) %>% 
  st_transform(., 27700) %>% 
  dplyr::select(cluster, geometry) %>% 
  rename(st_clusters = "cluster")

stPlot <- sfc_as_cols(stPlot, c("longitude", "latitude"))

hulls <- stPlot %>%
  group_by(st_clusters) %>%
  dplyr::mutate(hull = 1:n(),
                hull = factor(hull, chull(longitude, latitude)))%>%
  arrange(hull) %>% 
  filter(st_clusters >=1) # since dbcluster < 1 are noises

# plot
p <- autoplot.OpenStreetMap(basemap_bng) + 
  geom_point(data=hulls,
             aes(longitude,latitude,
                 colour=st_clusters,
                 fill=st_clusters),
             size=0.5, alpha=0.5) + 
  geom_polygon(data = hulls,
               aes(longitude,latitude,
                   group = st_clusters,
                   fill = st_clusters)) +
 theme(axis.text=element_text(size=7),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size= 7),
       legend.position = "right") +
  theme(plot.title=element_text(hjust = 0.05, face="bold"))+
  labs(caption = "Copyright OpenStreetMap contributors",
       x="Meter",y="Meter")

ggsave(file="output/stdbscan_map.png", p)
```

